---
title: "Soil_Omar"
author: "Omar Godinez"
date: "4/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
Necessary libraries to start working.
```{r}
library(tidyverse)
```

# Step 1: Load files
It is necessary to download the four files before start.
Here we load and assign a name to each one of them.
```{r}
SoilData_1 <- read_csv("data/Soil_1.csv")
SoilData_2 <- read_csv("data/Soil_2.csv")
SoilData_2 <- SoilData_2 %>% rename(vial = Core)
Soil_key_1 <- read_csv("data/Soil_Core_Key.csv")
Soil_key_1 <- Soil_key_1 %>% rename(Core = core)
Soil_key_2 <- read_csv("data/partitioned_key.csv")
```

# Step 2: Clean the data
Remove empty rows, join the data with the new keys, and create an index along with other new variables.
In simple words, prepare the data sets to start working.
```{r}
SoilData_1 <- SoilData_1[rowSums(is.na(SoilData_1)) != ncol(SoilData_1),]  #Remove rows with only NAs
SoilData_1 <- SoilData_1[-c(122),]                                         #Remove no measured week
SoilData_1 <- left_join(SoilData_1,Soil_key_1, by = "Core")                #Join both data
SoilData_1 <- subset(SoilData_1, select = -c(Treatment))                   #Remove the extra treatment
SoilData_1 <- SoilData_1 %>% rename(Treatment = treatment)                 #Rename Treatment
SoilData_1 <- SoilData_1 %>% mutate(CCO2 = (mmol_air*CO2_ppm*10^(-3)*12))               #Creates new variable
SoilData_1 <- unite(SoilData_1, Type.TRTMT, type, Treatment, sep = "_", remove = FALSE) #Uniting Type and treatment
SoilData_1 <- unite(SoilData_1, TempType, type, Treatment, sep = "_", remove = FALSE)
SoilData_1 <- SoilData_1 %>% group_by(TempType) %>%  mutate(index = row_number())     #Creating index number

SoilData_2 <- SoilData_2[rowSums(is.na(SoilData_2)) != ncol(SoilData_2),]            #Remove rows with only NAs
SoilData_2 <- filter(SoilData_2, Layer != "mineral extra test")                                              #Remove mineral extra tests
SoilData_2 <- SoilData_2[SoilData_2$vial != "1C extra", ]                            #Remove 1c extra
SoilData_2 <- left_join(SoilData_2,Soil_key_2, by = "vial")                          #Join both data
SoilData_2 <- SoilData_2 %>% mutate(CCO2 = (partition_mmol_air*CO2_ppm*10^(-3)*12))  #Creates new variable
SoilData_2 <- unite(SoilData_2, Vial.Layer, vial, Layer, sep = "_", remove = FALSE)  #Uniting Core and Layer
SoilData_2 <- SoilData_2 %>% group_by(Vial.Layer) %>%  mutate(index = row_number())  #Creating index number
SoilData_2 <- SoilData_2 %>% mutate(LayerCond = ifelse(Layer == "mineral", "Mineral", ifelse(Layer == "organic" | Layer == "organic bottom" | Layer == "organic top", "Organic", ifelse(Layer == "plant", "Plant", NA))))        #Groups all organic layers

```

# Step 3: Summaries
A look to the joined data.
```{r}
summary(SoilData_1)
summary(SoilData_2)
```

# Analyze CORE SOIL DATA

# Step 4: CCO2 analysis
```{r}
SoilData_1 %>% group_by(Treatment, type) %>% summarize(size = n(), mean = mean(CCO2), median = median(CCO2), st.dev = sd(CCO2), IQR = IQR(CCO2), min = min(CCO2), max = max(CCO2))
boxplot(SoilData_1$CCO2~SoilData_1$Treatment, main="CCO2 by treatment", xlab="Treatment", ylab = "CCO2")
boxplot(SoilData_1$CCO2~SoilData_1$Type.TRTMT, main="CCO2 by type/treatment", xlab="Type/Treatment", ylab = "CCO2")
ggplot(data = SoilData_1) +
  geom_smooth(mapping = aes(x = index, y = CCO2, color = Type.TRTMT )) + 
  ggtitle("CCO2's plot by type/treatment") +xlab("Index") + ylab("CCO2")
```

# Step 5: CO2_ppm analysis
```{r}
SoilData_1 %>% group_by(Treatment, type) %>% summarize(size = n(), mean = mean(CO2_ppm), median = median(CO2_ppm), st.dev = sd(CO2_ppm), IQR = IQR(CO2_ppm), min = min(CO2_ppm), max = max(CO2_ppm))
boxplot(SoilData_1$CO2_ppm~SoilData_1$Treatment, main="CO2_ppm by treatment", xlab="treatment", ylab = "CO2_ppm")
boxplot(SoilData_1$CO2_ppm~SoilData_1$Type.TRTMT, main="CO2_ppm by type/treatment", xlab="type/treatment", ylab = "CO2_ppm")
ggplot(data = SoilData_1) +
  geom_smooth(mapping = aes(x = index, y = CO2_ppm, color = Type.TRTMT )) + 
  ggtitle("CO2_ppm's plot by type/treatment") + xlab("Index") + ylab("CO2_ppm")
```

# Analyze Layer SOIL DATA

# Step 6: CCO2 analysis
```{r}
SoilData_2 %>% group_by(Treatment, Layer) %>% summarize(size = n(), mean = mean(CCO2), median = median(CCO2), st.dev = sd(CCO2), IQR = IQR(CCO2), min = min(CCO2), max = max(CCO2))
boxplot(SoilData_2$CCO2~SoilData_2$Layer, main="CCO2 by layer", xlab="Layer", ylab = "CCO2")
boxplot(SoilData_2$CCO2~SoilData_2$Vial.Layer, main="CCO2 by vial/layer", xlab="Vial/Layer", ylab = "CCO2")
boxplot(SoilData_2$CCO2~SoilData_2$LayerCond, main="CCO2 by layer group", xlab="Group", ylab = "CCO2")
ggplot(data = SoilData_2) +
  geom_smooth(mapping = aes(x = index, y = CCO2, color = LayerCond)) + 
  ggtitle("CCO2's plot by layer group") +xlab("Index") + ylab("CCO2")
ggplot(SoilData_2, aes(LayerCond, CCO2, fill = LayerCond)) + geom_bar(stat = 'identity') + 
  ggtitle("CCO2's generated") +xlab("Group") + ylab("CCO2")
ggplot(SoilData_2) + geom_histogram(mapping = aes(x = CCO2)) + facet_wrap(~LayerCond)
```

# Step 7: CO2_ppm analysis
```{r}
SoilData_2 %>% group_by(Treatment, Layer) %>% summarize(Size = n(), mean = mean(CO2_ppm), median = median(CO2_ppm), st.dev. = sd(CO2_ppm), IQR = IQR(CO2_ppm), min = min(CO2_ppm), max = max(CO2_ppm))
boxplot(SoilData_2$CO2_ppm~SoilData_2$Layer, main="CO2_ppm by layer", xlab="Layer", ylab = "CO2_ppm")
boxplot(SoilData_2$CO2_ppm~SoilData_2$Vial.Layer, main="CO2_ppm by vial/layer", xlab="Vial/Layer", ylab = "CO2_ppm")
boxplot(SoilData_2$CO2_ppm~SoilData_2$LayerCond, main="CO2_ppm by layer group", xlab="Group", ylab = "CO2_ppm")
ggplot(data = SoilData_2) +
  geom_smooth(mapping = aes(x = index, y = CO2_ppm, color = LayerCond)) + 
  ggtitle("CO2_ppm's plot by layer group") +xlab("Index") + ylab("CO2_ppm")
ggplot(SoilData_2, aes(LayerCond, CO2_ppm, fill = LayerCond)) + geom_bar(stat = 'identity') + 
  ggtitle("CO2_ppm's generated") +xlab("Group") + ylab("CO2_ppm")
ggplot(SoilData_2) + geom_histogram(mapping = aes(x = CO2_ppm)) + facet_wrap(~LayerCond)
```
